---
title: "R for EPP"
author: "Angelos Krypotos"
knit: slidify::knit2slides
mode: standalone
job: null
subtitle: Statistical Analyses with R
framework: io2012
widgets     : [mathjax angularjs]  
highlighter : highlight.js
hitheme     : tomorrow 
---

pre {
  font-size: 5px;
}

slide.smallcode pre {
  font-size: 5px;
}

<style>
  .title-slide {
    background-color:  #808080; /*; ; #CA9F9D*/
      /* background-image:url(http://goo.gl/EpXln); */
  }
</style>


<style>
.title-slide hgroup > h1, 
.title-slide hgroup > h2 {
  color: #F5F5F5 ;  /* ; #EF5150*/
}
</style>



--- 
## Goals of today's presentation

- Assignment 3 (Feedback)

- Analyses with R

- Decide next lecture

---
## Assignment 2

- Any comments?

---
## Basic descriptives

```{r}
library(psych, quietly = TRUE) # Load psych library
data(galton) # Load galton data
head(galton, n = 2) # Head of data
describe(galton) # Basic descriptives
```

---
## Correlations

```{r}
pairs.panels(galton, main="Galton's Parent child heights")  

```

---
## Correlations

```{r}
cor(galton$parent, galton$child)
cor.test(galton$parent, galton$child)
```

---
## Linear regression (1)

lm(DV~IV1)

```{r}
data(affect)
res <- lm(state1~state2, data = affect)
summary(res)
## Also run plot(res)
```

---

lm(DV~IV1+IV2): Only main effects


```{r}
res <- lm(state1~state2+NA2, data = affect)
summary(res)
```


---

lm(DV~IV1:IV2): Only interaction(s)

```{r}
res <- lm(state1~state2:NA2, data = affect)
summary(res)
```


---
lm(DV~IV1*IV2): All single effects and interaction(s)

```{r}
res <- lm(state1~state2*NA2, data = affect)
summary(res)
```

---
# t.tests
First, boxplots
```{r}
boxplot(neur~Study, data = affect)
```

---
# t.tests (Indepedent samples)

```{r}
t.test(neur~Study, data = affect)
```

---
# t.tests (Dependent samples)

```{r}
t.test(affect$PA1, affect$NA1, paired = TRUE)
```

---
# ANOVA

Multiple ways to do so

The _ez_ package is a good way to do ANOVAS easily

---
# ANOVA

```{r}
library(ez, quietly = TRUE)
data("ANT")
head(ANT, n = 3)
```

---
# ANOVA

```{r}
res <- ezANOVA(              ### The function
data = ANT[ANT$error==0,],    ### The data to be used
dv = rt,                   ### The dependent variable
wid = subnum,              ### The subject factor
within = .(cue,flank),     ### Within subject variable
between = group,           ### group variablew
type = 3                    ### Ensure that results are comparable to SPSS results
)
```


---
# ANOVA

```{r}
res$ANOVA
```


---
# ANOVA

```{r}
res$`Mauchly's Test for Sphericity`
```

---
# ANOVA

```{r}
res$`Sphericity Corrections`
```


---
# BAYES WILL TEAR US APART

- rbugs (R and WinBugs)


- rjags (R and JAGS)

- Rstan (R and STAN)

---
# BayesFactor

The BayesFactor package allows the easy computation of Bayes
factors for common statistical designs

BayesFactor is not suitable for parameter estimation

This [webpage](http://bayesfactorpcl.r-forge.r-project.org/) is absolutely amazing for further info 

---
## Bayesian t-test

```{r}
data(affect)
library(BayesFactor)
res = ttestBF(affect$neur[affect$Study == "maps"], affect$neur[affect$Study == "maps"], rscale = .707)
```

---
## Bayesian t-test

```{r}
res
1/res
```

---
## Bayesian regression

```{r}
lmBF(ext~Study, data = affect)
```

---
## Bayesian ANOVA

```{r}
res <- anovaBF(rt ~ cue *  group + subnum, whichRandom = "subnum",
  data = ANT, progress = FALSE)
```

---
## Bayesian ANOVA

```{r}
res
```

---
## Instructions on Assignment

- All assignments can be found [here](https://github.com/AngelosPsy/EPPR/tree/master/Assignments)
- You can hand in your assignments by sending me an html document, with, when
asked, runnable code.
- Go as far as possible with the assignment. If you cannot complete it, no problem.
- I AM AVAILABLE FOR QUESTIONS!!!!!!
- Note: make sure that you remove the "eval = T", wherever you see it.
- Note: For writing text you should start each line with "#'" (without the quotation marks)

---

## The end
